---
format: 
  revealjs:
    pagetitle: "Area Data III"
    menu: false
    footer: "2026 Zehui Yin"
---

## $\LaTeX$ annual update

If you installed $\LaTeX$ with the `tinytex` package, you may want to update it to the latest version. Around the same time each year, TeX Live releases a new version of the distribution. If you have an older version of TeX Live, you may encounter errors when needing to install new $\LaTeX$ packages.

You can do that by running the following code in the console.

```{r, echo=TRUE, eval=FALSE}
tinytex::uninstall_tinytex()

tinytex::install_tinytex()
```

## Linear regression

:::{style="font-size: 0.8em;"}

The theoretical regression model:

$$
\underbrace{y_i}_{\text{Dependent Variable}} = 
  \underbrace{\beta_0}_{\text{Intercept}} + \underbrace{\beta_1x_{i1} + \beta_2x_{i2} + \cdots + \beta_jx_{ij}}_{\text{Independent Variables}} + \underbrace{\epsilon_i}_{\text{Random Error}}
$$

An estimated regression equation:

$$
\hat{y_i} = \hat{\beta_0} + \hat{\beta_1}x_{i1} + \hat{\beta_2}x_{i2} + \cdots + \hat{\beta_j}x_{ij}
$$

The loss function minimized with the OLS procedure:

$$
\sum e_i^2 = \sum (y_i - \hat{y_i})^2 = \sum (y_i - \hat{\beta_0} - \hat{\beta_1}x_{i1} - \hat{\beta_2}x_{i2} - \cdots - \hat{\beta_j}x_{ij})^2
$$
:::

## Trend surface

General form:

$$
y_i = f(lat_i,long_i) + \epsilon_i
$$

First degree:

$$
y_i = \beta_0 + \beta_1lat_i +\beta_2long_i + \epsilon_i
$$

Second degree:

$$
y_i = \beta_0 + \beta_1lat_i^2 +\beta_2lat_i + \\
\beta_3lat_i \cdot long_i +\beta_4long_i + \beta_5long_i^2 + \epsilon_i
$$

## Spatially-varying Coefficients

:::{style="font-size: 0.8em;"}

**Expansion Method:**

Use the trend surface function in place of a constant coefficient:

$$
y_i = f_{(lat_i, long_i)}^1 + f_{(lat_i, long_i)}^2 x_{i1} + \epsilon_i
$$

$$
\begin{array}{l}
f_{(lat_i, long_i)}^1 = \beta_{01} +\beta_{02}lat_i + \beta_{03}long_i\\
f_{(lat_i, long_i)}^2 = \beta_{11} +\beta_{12}lat_i + \beta_{13}long_i
\end{array}
$$

**Geographically Weighted Regression:**

Instead of fitting a regression model to all the observations, it fits to a subset of them each time based on the bandwidth defined.

It is like a spatial moving average; instead of calculating an average, we are fitting a regression model to each window.
:::

## Spatial error model

$$
y_i = \beta_0 + \beta_1x_{i1} + \beta_2x_{i2} + \cdots + \beta_jx_{ij} + \underbrace{\epsilon_i}_{\text{Random Error}}
$$

The error term is no longer independent of each other but follows an autoregressive form:

$$
\epsilon_i = \lambda \sum_k w_{ik}\epsilon_k + u_i
$$

Noting that the $u_i$ is the pure random component now.

## Capstone activity expectation

For lab activity submission, you need to at least conduct **descriptive statistics**.

Load the data and use the `summary` method or provide some descriptive plots like histograms.

```{r}
df <- mtcars[, 1:4]
```

```{r, echo=TRUE}
summary(df)
```

## Activities for today

- We will work on the following chapter from the textbook:
  - Chapter [28](https://paezha.github.io/spatial-analysis-r/activity-13-area-data-v.html): Activity 13: Area Data V
  - Chapter [30](https://paezha.github.io/spatial-analysis-r/activity-14-area-data-vi.html): Activity 14: Area Data VI
- The hard deadline is **Tuesday**, **March 10**.
