---
format: 
  revealjs:
    pagetitle: "Point Pattern II"
    menu: false
    footer: "2025 Zehui Yin"
---

## Activity feedback

- Do **not** submit last week's activity.
  - Double-check what you submitted on Avenue!
- We have **two** activities each week, and you need to submit both in each submission.
- A small $p$-value ($<0.05$) in the $\chi^2$ test indicates that the point pattern is not random and shows spatial dependence.

## Mathematical notation remainder

Summation notation ($\sum$):

$$
\sum_{i=1}^Qx_i=x_1 + x_2 + \cdots + x_Q
$$

For all notation ($\forall$):

$$
x_k = 1 \hspace{0.2cm} \forall k\rightarrow x=1
$$

## $\chi^2$ distribution

Strictly non-negative, right-skewed, and non-symmetric.

```{r}
library(ggplot2)
library(dplyr)
set.seed(437988)

rchisq(10000, df = 10) |> as.data.frame() |>
  `colnames<-`("chi") |>
  ggplot() + 
  geom_histogram(aes(x = chi),
                 binwidth = 2) +
  labs(x = "", y = "Frequency") +
  theme_bw()
```

## Cumulative distribution function

$F_X(x)=P(X \leq x)$

```{r}
rchisq(10000, df = 10) |> 
  ecdf() |> 
  quantile(probs = seq(0, 0.99, 0.025)) |>
  as.data.frame() |>
  `colnames<-`("chi") |>
  mutate(prob = seq(0, 0.99, 0.025)) |>
  ggplot() + 
  geom_point(aes(x = chi, y = prob)) +
  geom_line(aes(x = chi, y = prob)) +
  labs(x = "Chi-squared", y = "Proportion") +
  theme_bw()
```

## $\hat{G}$-function

$$
\hat{G}(r) = \frac{1}{n(\mathbf
x)}\sum_{i}\mathbf 1\{d_i \le r\}
$$

The $\hat{G}$ function represents the number of elements in the set of distances up to some threshold $r$, normalized by the total number of points $n$ in point pattern $\mathbf{x}$.

## Pipe operator in R

The `|>` is the native pipe operator introduced in R 4.1.0.

The old pipe operator is `%>%` from `magrittr` package.

It allows chained expressions in place of nested expressions to improve code readability.

```{r, echo=TRUE}
v <- c(1, 2, 3, 4, 5)
sum(sqrt(v))
```

```{r, echo=TRUE}
v |> 
  sqrt() |>
  sum()
```

Do not confuse it with `+` in `ggplot2`; they are **not** the same thing.

## Packages we use today

Load the following three packages.

```{r, echo=TRUE, eval=FALSE}
library(isdas)
library(sf)
library(tidyverse)
library(spatstat)
```

If you have trouble restoring the reproducible environment, you need to manually install the packages first.

```{r, echo=TRUE, eval=FALSE}
install.packages("remotes")
remotes::install_github("paezha/isdas")

install.packages("sf")

install.packages("tidyverse")

install.packages("spatstat")
```

## Activities for today

- We will work on the following chapter from the textbook:
  - Chapter [12](https://paezha.github.io/spatial-analysis-r/activity-5-point-pattern-analysis-ii.html): Activity 5: Point Pattern Analysis II
  - Chapter [14](https://paezha.github.io/spatial-analysis-r/activity-6-point-pattern-analysis-iii.html): Activity 6: Point Pattern Analysis III
- The hard deadline is **Friday**, **February 7**.

## Reference

- <https://www.geo.fu-berlin.de/en/v/soga-r/Advances-statistics/Spatial-Point-Patterns/Analysis-of-Spatial-Point-Patterns/Interactions-in-Point-Pattern-Analysis/index.html>
